**The status of the microbial census: an update**

### 2015-11-30 Schloss_Census2_mBio_2015.Rmd
Incorporate culturability analysis

```{r culture}
	b_is_cult <- (!is.na(bact$strain) | !is.na(bact$isolate)) & !grepl("\\.Unc", rownames(bact))
	a_is_cult <- (!is.na(arch$strain) | !is.na(arch$isolate)) & !grepl("\\.Unc", rownames(arch))

	b_ncult_full <- sum(b_is_cult)
	a_ncult_full <- sum(a_is_cult)

	b_fract_cult <- 100 * b_ncult_full / nrow(bact)
	a_fract_cult <- 100 * a_ncult_full / nrow(arch)

	b_ncult_2005 <- sum(b_is_cult & bact_year_deposited <= 2005, na.rm=T)
	a_ncult_2005 <- sum(a_is_cult & arch_year_deposited <= 2005, na.rm=T)

	b_ncult_delta <- b_ncult_full - b_ncult_2005

	b_before_2005 <- sum(bact_year_deposited <= 2005, na.rm=T)
	b_since_2005 <- sum(bact_year_deposited > 2005, na.rm=T)

	b_slow_cult <- (b_since_2005/b_before_2005) / (b_ncult_delta/b_ncult_2005)

	a_ncult_delta <- a_ncult_full - a_ncult_2005

	a_before_2005 <- sum(arch_year_deposited <= 2005, na.rm=T)
	a_since_2005 <- sum(arch_year_deposited > 2005, na.rm=T)

	a_slow_cult <- (a_since_2005/a_before_2005) / (a_ncult_delta/a_ncult_2005)

	bact_phylum <- gsub("^[^;]*;([^;]*);.*", "\\1", bact$taxonomy)
	bact_phylum[bact_phylum=="Euryarchaeota"] <- NA
	bact_phylum[bact_phylum=="aquifer1"] <- NA

	b_is_cult[rownames(bact) == 'AJ224540.MxeCultu'] <- FALSE
#	aquifer1 - scrap
#	bact[bact_phylum=="Candidate_division_OP3" & !is.na(bact$isolate) & !is.na(bact_phylum),] - AJ224540.MxeCultu

#	bact[bact_phylum=="CKC4" & !is.na(bact$strain) & !is.na(bact_phylum),] -> 2 - 1 is a genome, 1 is an isolate from same research group
#	bact[bact_phylum=="PAUC34f" & (!is.na(bact$isolate) | !is.na(bact$strain)),] - legit
#	bact[bact_phylum=="TM6" & (!is.na(bact$isolate) | !is.na(bact$strain)) & !grepl("\\.Unc", rownames(bact)),] - legit

	b_phylum_cult_table <- table(bact_phylum, b_is_cult)
	b_phylum_cult_rate <- b_phylum_cult_table[,"TRUE"] / apply(b_phylum_cult_table, 1, sum)

	arch_phylum <- gsub("^[^;]*;([^;]*);.*", "\\1", arch$taxonomy)
	arch_phylum[arch_phylum=="Archaea;"] <- "Unclassified"
	a_phylum_cult_table <- table(arch_phylum, a_is_cult)
	a_phylum_cult_rate <- a_phylum_cult_table[,"TRUE"] / apply(a_phylum_cult_table, 1, sum)

	b_cult_otus <- unique(bact$otu[b_is_cult])
	b_in_cult_otu <- bact$otu %in% b_cult_otus

	n_all_otus_phylum <- aggregate(bact$otu, by=list(bact_phylum), function(x){length(unique(x))})
	n_cult_otus_phylum <- aggregate(bact$otu[b_in_cult_otu], by=list(bact_phylum[b_in_cult_otu]), function(x){length(unique(x))})

	b_otus_phylum <- data.frame(matrix(rep(0, nrow(n_all_otus_phylum)*2), ncol=2, dimnames=list(n_all_otus_phylum$Group.1, c('cultured', 'total'))))

	b_otus_phylum[n_all_otus_phylum$Group.1,"total"] <- n_all_otus_phylum$x
	b_otus_phylum[n_cult_otus_phylum$Group.1,"cultured"] <- n_cult_otus_phylum$x

	a_cult_otus <- unique(arch$otu[a_is_cult])
	a_in_cult_otu <- arch$otu %in% a_cult_otus

	n_all_otus_phylum <- aggregate(arch$otu, by=list(arch_phylum), function(x){length(unique(x))})
	n_cult_otus_phylum <- aggregate(arch$otu[a_in_cult_otu], by=list(arch_phylum[a_in_cult_otu]), function(x){length(unique(x))})

	a_otus_phylum <- data.frame(matrix(rep(0, nrow(n_all_otus_phylum)*2), ncol=2, dimnames=list(n_all_otus_phylum$Group.1, c('cultured', 'total'))))

	a_otus_phylum[n_all_otus_phylum$Group.1,"total"] <- n_all_otus_phylum$x
	a_otus_phylum[n_cult_otus_phylum$Group.1,"cultured"] <- n_cult_otus_phylum$x

	#not sure what this is actually getting...
	b_n1_cult_otus <- 100*sum(table(bact$otu[b_in_cult_otu])==1)/length(table(bact$otu[b_in_cult_otu]))
	a_n1_cult_otus <- 100*sum(table(arch$otu[a_in_cult_otu])==1)/length(table(arch$otu[a_in_cult_otu]))
```

***The cultured census***
In the 2004 bacterial census there was great concern that although culture-independent methods were significantly enhancing our knowledge of microbial life, there were numerous bacterial phyla with no or only a few cultured representatives. To update this assessment, we identified those sequences that came from cultured and uncultured organisms. Overall, `r b_fract_cult`% of bacterial sequences and `r a_fract_cult`% of archaeal sequences have come from isolated organisms. Comparing the fraction of sequences deposited during and before 2006 from isolates to those collected after 2006, we found that culturing rates lag by `r  b_slow_cult` and `r a_slow_cult`-fold for bacteria and archaea, respectively. Among the `r length(b_phylum_cult_rate)` bacterial phyla, `r sum(b_phylum_cult_rate == 0)` have no cultured representatives and `r sum(b_phylum_cult_rate == 0)` of the `r length(a_phylum_cult_rate)` archaeal phyla have cultured representatives. Of the phyla with at least one cultured representative, the median percentage of sequences coming from a culture was only `r median(b_phylum_cult_rate[b_phylum_cult_rate!=0])*100`% for the bacterial phyla and `r median(a_phylum_cult_rate[a_phylum_cult_rate!=0])*100`% for the archaeal phyla (Figure XXXXXX). When we asked what percentage of OTUs had at least one cultured respresentative we found that `r 100*sum(b_otus_phylum$cultured)/sum(b_otus_phylum$total)`% of the `r sum(b_otus_phylum$total)` bacterial OTUs and `r 100*sum(a_otus_phylum$cultured)/sum(a_otus_phylum$total)`% of the `r sum(a_otus_phylum$total)` archeael OTUs had at least one cultured representative. Of those OTUs that had a sequence from a culture, `r b_n1_cult_otus`
and `r a_n1_cult_otus`% of them only contained sequences from cultures. These values suggest that bacterial cultivation efforts are biased towards

* need a plot showing % cultured sequences and % cultured OTUs
	* fraction of sequences from cultured / uncultured bugs (2014) by phylum
		- b_phylum_cult_rate
		- a_phylum_cult_rate
	* % of OTUs with a cultured representative by phylum and in aggregate
		- b_otus_phylum
		- a_otus_phylum

>>>>

```{r culture}
b_cult_time <- read.table(file='data/process/bacteria.cultured_by_time_counts.tsv', header=T)
a_cult_time <- read.table(file='data/process/archaea.cultured_by_time_counts.tsv', header=T)

b_percent_cult_nseqs_all <- 100*sum(b_cult_time[,"all_cult_nseqs"])/sum(b_cult_time[,"all_total_nseqs"])
a_percent_cult_nseqs_all <- 100*sum(a_cult_time[,"all_cult_nseqs"])/sum(a_cult_time[,"all_total_nseqs"])

b_shift_cult_nseqs_all <- 1/((sum(b_cult_time[,"post_cult_nseqs"])/sum(b_cult_time[,"pre_cult_nseqs"])) / (sum(b_cult_time[,"post_total_nseqs"])/sum(b_cult_time[,"pre_total_nseqs"])))

a_shift_cult_nseqs_all <- 1/((sum(a_cult_time[,"post_cult_nseqs"])/sum(a_cult_time[,"pre_cult_nseqs"])) / (sum(a_cult_time[,"post_total_nseqs"])/sum(a_cult_time[,"pre_total_nseqs"])))

b_rate_cult_nseqs_phyulm <- b_cult_time[, 'all_cult_nseqs'] / b_cult_time[, 'all_total_nseqs']
a_rate_cult_nseqs_phyulm <- a_cult_time[, 'all_cult_nseqs'] / a_cult_time[, 'all_total_nseqs']

b_percent_cult_otu_all <- 100*sum(b_cult_time[,'all_cult_sobs'])/sum(b_cult_time[,'all_total_sobs'])
a_percent_cult_otu_all <- 100*sum(a_cult_time[,'all_cult_sobs'])/sum(a_cult_time[,'all_total_sobs'])
```

***The cultured census***
In the 2004 bacterial census there was great concern that although culture-independent methods were significantly enhancing our knowledge of microbial life, there were numerous bacterial phyla with no or only a few cultured representatives. To update this assessment, we identified those sequences that came from cultured and uncultured organisms. Overall, `r b_percent_cult_nseqs_all`% of bacterial sequences and `r a_percent_cult_nseqs_all`% of archaeal sequences have come from isolated organisms. Comparing the fraction of sequences deposited during and before 2006 from isolates to those collected after 2006, we found that culturing rates lag by `r  b_shift_cult_nseqs_all` and `r a_shift_cult_nseqs_all`-fold for bacteria and archaea, respectively. Among the `r nrow(b_cult_time)` bacterial phyla, `r sum(b_cult_time$all_cult_nseqs == 0)` have no cultured representatives and `r nrow(a_cult_time)` of the `r sum(a_cult_time$all_cult_nseqs == 0)` archaeal phyla have cultured representatives. This lag is likely due to the differences in throughput of culture-dependent and -independent approaches. Of the phyla with at least one cultured representative, the median percentage of sequences coming from a culture was only `r median(b_rate_cult_nseqs_phyulm[b_rate_cult_nseqs_phyulm!=0])*100`% for the bacterial phyla and `r median(a_rate_cult_nseqs_phyulm[a_rate_cult_nseqs_phyulm!=0])*100`% for the archaeal phyla (Figure phylum_effort.pdf). So, even though many phyla have cultured representatives, there is still a skew in the representation of most phyla found in cultivation efforts. Considering the possibility that large culture-independent sequencing efforts may only be re-sequencing organisms that already exist in culture, we asked what percentage of OTUs had at least one cultured representative. We found that `r b_percent_cult_otu_all`% of the `r sum(b_cult_time[,'all_total_sobs'])` bacterial OTUs and `r a_percent_cult_otu_all`% of the `r sum(a_cult_time[,'all_total_sobs'])` archeael OTUs had at least one cultured representative. Considering that the percentage of sequences from bacterial cultures is greater than the percentage of OTUs from bacterial cultures, we suspect that bacterial cultivation efforts are largely resampling the same diversity relative to culture-independent approaches. In contrast, the relationship is reversed among the archaea indicating that archaeal cultivation efforts continue to sample greater biodiversity than culture-independent approaches.

###### here #######
